---
title: "Formatted Spreadsheet Generator"
author: "Jesse Cambon"

date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
    toc: true
---

We will use the classic gapminder dataset to create a formatted spreadsheet report using the openxlsx package. This code allows us to create the spreadsheet report directly from code instead of using manual copy and pasting and formatting. This is especially conveniant for recurring reports.

To view the resulting spreadsheet report, [click here](gapminder_summary_report.xlsx).

## Setup

```{r,results=F,warning=F,message=F}
output_file <- 'gapminder_summary_report.xlsx'


library(gapminder)
library(tidyverse)
library(openxlsx)
library(knitr)
library(kableExtra)

gap_continent_summ <- gapminder %>% 
  filter(year==2007) %>% select(-year) %>%
  group_by(continent) %>%
  summarize(lifeExp=weighted.mean(lifeExp,pop),
            gdpPercap=weighted.mean(gdpPercap,pop),
            pop=sum(as.numeric(pop))) %>%
  ungroup() %>%
  mutate(header_level=1)

# Grab the top 5 most populuous countries from each continent
gap_top_countries <- gapminder %>%
  filter(year==2007) %>% select(-year) %>%
  group_by(continent) %>%
  arrange(desc(pop)) %>% slice(1:5) %>%
  ungroup() %>% 
  mutate(indent_level=1)

gap_combi <- bind_rows(gap_continent_summ,gap_top_countries) %>%
  arrange(continent,desc(pop)) %>%
  # combine country and continent fields for stacking
  mutate(category=coalesce(as.character(country),as.character(continent))) %>%
  select(category,lifeExp,gdpPercap,pop,everything()) %>%
  select(-country,-continent) %>%
  rename(`Life Expectancy` = lifeExp, `GDP Per Capita`=gdpPercap, Population=pop)

```

```{r tablepreview,results='asis'}
kable(gap_combi %>% select(-header_level,-indent_level) %>% rename(`Continent/Country`=category),format='markdown',
      format.args=list(big.mark=','),digits=c(0,1,0,0)) %>%
  kable_styling(bootstrap_options = c("striped",'border'))
```


## Create Spreadsheet

Now we will create the formatted spreadsheet

Comma, rounded (decimal), and percent columns can be entered either as column position or as
a column name. You can pass 'indent_level' and 'header_level' columns to the create_worksheet
function for formatting purposes. These columns will not be shown in the resulting spreadsheet.

```{r}
## Define the function for creating a worksheet
create_worksheet <- function(wb,sheet_name,data,title1_text="",comma_cols=NULL,percent_cols=NULL,
                             round_cols=NULL) {
  
  # Extract the indent values if they are provided 
  # and then remove from dataframe
  if (("indent_level") %in% colnames(data)) {

    indent_1_rows <- which(1 == data$indent_level)
    #print(indent_1_rows)
    indent_2_rows <- which(2 == data$indent_level)
    data <- data %>% select(-indent_level)
  }
  
  # Extract the header level values if they are provided 
  # and then remove from dataframe
  if (("header_level") %in% colnames(data)) {

    header_1_rows <- which(1 == data$header_level)
    header_2_rows <- which(2 == data$header_level)
    data <- data %>% select(-header_level)
  }
  
  # If percent and column columns were provided in character
  # (variable name) format then extract the column numbers
  if (class(percent_cols)=='character') {
    percent_cols = which(colnames(data) %in% percent_cols)
  }
  if (class(comma_cols)=='character') {
    comma_cols = which(colnames(data) %in% comma_cols)
  }
  
  if (class(round_cols)=='character') {
    round_cols = which(colnames(data) %in% round_cols)
  }

  # delete column header
  colnames(data) <- str_replace_all(colnames(data),
            c('category' = ''))

  
  # a thick vertical border will be drawn to the right of these columns
  sep_borders <- c()
    
  ## row on which column headers are (data table begins on this row)
  table_header_row = 3
  sheet_header_row = table_header_row - 1 # Overall sheet title
  # grouping_header_row = table_header_row - 1 # titles that group columns
  
  table_width = ncol(data) # number of columns
  table_last_row = table_header_row + nrow(data) # last row of data in the table

  ##############
  
  # Create worksheet
  addWorksheet(wb, sheet_name)
  
  # Style for sheet title
  table_header_style <- createStyle(halign='center',valign='center',textDecoration="Bold")
  
  # Style for column headers
  column_header_style <- createStyle(halign='center',valign='center',textDecoration="Bold",
                                     fgFill='#D9D9D9')
  
  ### Titles 
  # Merge cells for titles
  mergeCells(wb, sheet_name, cols = 2:(table_width+1), rows = sheet_header_row) # Overall Header
  
  # Write titles to Worksheet
  writeData(wb,sheet_name,title1_text,startCol=2,startRow=sheet_header_row)

  ## Style titles
  addStyle(wb,sheet_name,style=table_header_style,cols=1:table_width+1,rows=sheet_header_row,gridExpand = T,stack=T)

  # Write main data table to worksheet
  writeData(wb,sheet_name,data,startRow = table_header_row,startCol=2)
  
  ### Table styling
  # Draw borders for all cells in table
  addStyle(wb,sheet_name, style=createStyle(border='TopBottomLeftRight'), cols=1:table_width+1,
           rows=table_header_row:table_last_row,gridExpand = T)
  
  ## Format specified columns as percent, comma, or rounded (decimal) format
  if (!is.null(comma_cols)) {
  ## Apply Comma format
  addStyle(wb,sheet_name, style=createStyle(numFmt='COMMA'), cols=comma_cols+1,
           rows=table_header_row+1:table_last_row-table_header_row,gridExpand = T,stack=T)
  }
  if (!is.null(percent_cols)) {
  ## Apply Percent format
  addStyle(wb,sheet_name, style=createStyle(numFmt='PERCENTAGE'), cols=percent_cols+1,
           rows=table_header_row+1:table_last_row-table_header_row,gridExpand = T,stack=T)
  }
  
  if (!is.null(round_cols)) {
  ## Apply Rounded format
  addStyle(wb,sheet_name, style=createStyle(numFmt="0.0"), cols=round_cols+1,
           rows=table_header_row+1:table_last_row-table_header_row,gridExpand = T,stack=T)
  }

  ## Apply header formats to specifed rows
  if (exists("header_1_rows")) {
    # Bold and Center
  addStyle(wb,sheet_name,style=createStyle(textDecoration = "Bold"),
           cols=c(2),rows=table_header_row+header_1_rows,stack=T)
  # Highlight
  addStyle(wb,sheet_name,style=createStyle(fgFill='#BFBFBF'),
           cols=1:table_width+1,rows=table_header_row+header_1_rows,stack=T,gridExpand = T)
  }
  if (exists("header_2_rows")) {
  addStyle(wb,sheet_name,style=createStyle(textDecoration="bold"),cols=c(2),
         rows=table_header_row+header_2_rows,stack=T)
    
  addStyle(wb,sheet_name,style=createStyle(fgFill='#D9D9D9'),
           cols=1:table_width+1,rows=table_header_row+header_2_rows,stack=T,gridExpand = T)  
  }
  
  ## Indent specifed rows (just the first column) 
  if (exists("indent_1_rows")) {
  addStyle(wb,sheet_name, style=createStyle(indent=1), cols=c(2),
           rows=table_header_row+indent_1_rows,gridExpand = T,stack=T)
  }
  if (exists("indent_2_rows")) {
  addStyle(wb,sheet_name, style=createStyle(indent=2), cols=c(2),
           rows=table_header_row+indent_2_rows,gridExpand = T,stack=T)
  }

  # Column headers formatting
  addStyle(wb,sheet_name,style=column_header_style,
    cols=1:table_width+1, rows=table_header_row,
         gridExpand = T,stack=T)
  
  ## Draw thick borders surrounding table and also around
  # select columns within the table
  addStyle(wb,sheet_name,style=createStyle(border='Top',borderStyle='medium'),
           cols=2:(table_width+1),rows=c(table_header_row,table_last_row+1),stack=T,gridExpand = T )
  addStyle(wb,sheet_name,style=createStyle(border='Left',borderStyle='medium'),
           cols=c(2,table_width+2,sep_borders+1),rows=table_header_row:table_last_row,stack=T,gridExpand = T )
  
  setColWidths(wb,sheet= sheet_name,cols=2:(table_width+1),widths = c(35,rep(15,table_width-1)))
}

## Create the spreadsheet

wb <- createWorkbook()
create_worksheet(wb,"Summary",gap_combi,title1_text="Gapminder Data by Continent",comma_cols=c('Population'),
                 round_cols=c('Life Expectancy','GDP Per Capita'))
create_worksheet(wb,"All Countries",gapminder %>% filter(year==2007) %>% select(-year),title1_text="All Gapminder Countries",comma_cols=c('pop'),
                 round_cols=c('gdpPercap','lifeExp'))

saveWorkbook(wb, file = output_file, overwrite = T)
```

