---
title: "Modeling Workflow"
author: "Jesse Cambon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
    toc: true
---

Demonstrate model workflows with tidyverse, modelr, and broom.

## References
* http://r4ds.had.co.nz/many-models.html 

## Setup

```{r setup,warning=F,message=F}
library(tidyverse)
library(gapminder)
library(broom)
library(modelr)
library(knitr)
library(kableExtra)

```

## Exploration


These graphs show why log transforming GDP per Capita makes it correlate more linearly to our response variable, life expectancy. Log transformations are often useful for highly skewed variables in regression.

```{r explore}
ggplot(data=gapminder,
          aes(x = gdpPercap, y = lifeExp, color = continent,group=1)) +
geom_point(alpha=0.7) +
theme_bw() +
geom_smooth() +
theme(legend.position='top',
  plot.title = element_text(lineheight=1, face="bold",hjust = 0.5)) + 
guides(color=guide_legend(override.aes = list(size=2.5))) 

ggplot(data=gapminder,
          aes(x = log(gdpPercap), y = lifeExp, color = continent,group=1)) +
geom_point(alpha=0.7) +
theme_bw() +
geom_smooth() +
theme(legend.position='top',
  plot.title = element_text(lineheight=1, face="bold",hjust = 0.5)) + 
guides(color=guide_legend(override.aes = list(size=2.5))) 


```



## Grouped Models

```{r models}

# One model per continent
models <- gapminder %>%
  group_by(continent) %>%
  do(fit=lm(lifeExp ~ (log(gdpPercap)*log(pop)) + year, data=.)) 

stats <- glance(models,fit) %>%
  arrange(desc(r.squared))

coefficients <- tidy(models,fit) %>%
  filter(term != '(Intercept)') %>%
  arrange(continent,desc(p.value))

fit_check <- augment(models,fit)
```


```{r plot}

ggplot(data=fit_check,
          aes(x = .fitted, y = .resid, color = continent,group=1)) +
geom_point(alpha=0.8) +
facet_grid(~continent) +
ggtitle('Fitted vs. Residual Check') +
theme_bw() +
geom_hline(yintercept=0,color='blue') + # horizontal line at 0 residual
theme(legend.position='none',
  plot.title = element_text(lineheight=1, face="bold",hjust = 0.5)) + 
guides(color=guide_legend(override.aes = list(size=2.5))) +
xlab('Fitted') +
ylab('Residual')

ggplot(data=fit_check,
          aes(.resid)) +
geom_histogram(aes(fill=continent)) +
facet_grid(~continent) +
ggtitle('Residual Distribution') +
theme_bw() +
scale_y_continuous(expand = c(0,0,0.05,0)) + 
theme(legend.position='none',
  plot.title = element_text(lineheight=1, face="bold",hjust = 0.5)) + 
guides(color=guide_legend(override.aes = list(size=2.5))) +
xlab('Residual') +
ylab('Count')
```



```{r modeldisplay,results='asis',warning=F}
kable(stats,format='markdown',digits=2) %>%
  kable_styling(bootstrap_options = c("striped",'border'))

kable(coefficients,format='markdown',digits=4) %>%
  kable_styling(bootstrap_options = c("striped",'border'))
```



## Nested Models

Now we create a similar model with nesting

```{r}
my_model <- function(df) {
  lm(lifeExp ~ (log(gdpPercap)*log(pop)) + year, data= df)
}

# Nest models by continent 

nested_models <- gapminder %>% 
  group_by(continent,country) %>% 
  nest() %>%
  # fit models
  mutate(fit = map(data, my_model)) %>%
  # calculate residuals
  mutate(augment = map(fit, augment),
    stats = map(fit,glance),
    terms = map(fit,tidy)) %>%
  ungroup()

# Dataset with predictions and residuals
model_fit <- nested_models %>% unnest(augment)

nest_stats <- nested_models %>%
  unnest(stats,.drop=TRUE) %>%
  arrange(desc(r.squared)) 

nest_coefficients <- nested_models %>%
  unnest(terms,.drop=TRUE) %>%
  filter(term != '(Intercept)') %>%
  arrange(continent,country,desc(p.value))

most_important_vars <- nest_coefficients %>%
  group_by(country) %>% 
  slice(1)

summ_imp_vars <- most_important_vars %>%
  group_by(continent) %>%
  count(term) %>%
  arrange(continent,desc(n))
```


