---
title: "Bayesian Modeling"
author: "Jesse Cambon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
    toc: true
---

```{r knit-settings, include=FALSE}
library(here)
source(here("rmd_config.R"))
```

References:
* http://appliedpredictivemodeling.com/data
* http://faculty.marshall.usc.edu/gareth-james/ISL/data.html

## Setup

```{r,message=F,warning=F}
#library(AppliedPredictiveModeling) # datasets
library(ISLR) # datasets
library(skimr)
library(tidyverse)
library(wesanderson)
library(rstanarm)
library(bayestestR)
library(insight)
library(bayesplot)
library(broom)
library(rsample)
library(knitr)
library(jcolors)
library(patchwork)
library(ggrepel)

num_cores <-  parallel::detectCores()
options(mc.cores = num_cores)

set.seed(42) # for reproducibility

# C/V split
split <- initial_split(Carseats, prop = 1/2)
carseat_train <- training(split) %>% as_tibble()
carseat_test  <- testing(split) %>% as_tibble()

carseat_train_small <- carseat_train %>% sample_n(50)
carseat_train_tiny <- carseat_train %>% sample_n(10)
```

Fit models

```{r}
lm_model <- lm(Sales ~ Advertising + Price, data = carseat_train)
stan_model <- stan_glm(Sales ~ Advertising + Price, data = carseat_train)
stan_model_small <- stan_glm(Sales ~ Advertising + Price, data = carseat_train_small)
stan_model_tiny <- stan_glm(Sales ~ Advertising + Price, data = carseat_train_tiny)
```

```{r}
tidy(lm_model) %>% kable()
tidy(stan_model) %>% kable()
```

```{r}
prior_summary(stan_model)
```

```{r}
prior_summary(stan_model_small)
```

```{r}
prior_summary(stan_model_tiny)
```

Draw from the prior and posterior distributions

```{r}
# Function for simulating prior and posterior distributions from stan model
sim_post_prior <- function(model) {
  # Simulate prior with bayestestR package
  prior <- simulate_prior(model) %>%
  pivot_longer(everything(),names_to='Parameter')

  # Simulate Posterior with insight package
  posterior <- get_parameters(model,iterations=10000) %>% 
  pivot_longer(everything(),names_to='Parameter')

  # Combine into one dataset
  combined <- prior %>% mutate(Distribution='Prior') %>% 
  bind_rows(posterior %>% mutate(Distribution='Posterior'))
  
  return(combined)
}

prior_posterior <- sim_post_prior(stan_model)
prior_posterior_small <- sim_post_prior(stan_model_small)
prior_posterior_tiny <- sim_post_prior(stan_model_tiny)

```

Plot our parameter prior and posterior distributions

```{r}

# Find the x,y coordinates for peak density in a sample
find_peak_density <- function(x_sample) {
# need the unlist() call to read from nested tibble data
density_x <- density(x_sample)
# Find coordinates for peak density
x_max <- density_x$x[which.max(density_x$y)]
y_max <- max(density_x$y)

return(tibble(x=x_max,y=y_max))
}

# Function for plotting 
plot_parameters <- function(distribution_sample,train_data) {
    
  # data to plot - exclude intercept term
  plot_data <- distribution_sample %>% filter(!str_detect(Parameter,'Intercept'))
    
  # Points for labeling max density 
  # based loosely on: https://stackoverflow.com/questions/56520287/how-to-add-label-to-each-geom-density-line)
 density_coordinates <- plot_data %>% 
  group_by(Distribution,Parameter) %>%
  do(find_peak_density(.$value))
    
  ggplot(data=plot_data,
         aes(x=value,fill=Parameter)) +
    facet_wrap(~fct_rev(Distribution),scales='free') +
    theme_minimal() +
    scale_y_continuous(expand =c(0,0,0.15,0)) + # add spacing for labels
    geom_vline(xintercept=0,color='red',size=0.25,linetype='dashed') +
    geom_point(data=density_coordinates, aes(x=x, y=y),show.legend = F) +
    geom_text_repel(data=density_coordinates, aes(label=round(x,2),x=x, y=y),
                     force=1.5,size=4,show.legend = F) +
    theme(legend.position='top',
          legend.title=element_blank(),
          plot.title = element_text(hjust = 0.5)) +
    geom_density(alpha=0.4,size=0.05) + ggtitle(str_c('n = ',as.character(nrow(train_data)))) +
    xlab('') + ylab('') + scale_fill_jcolors('pal6') + 
    guides(color = guide_legend(reverse=T))
}
```

Compare parameter distributions by sample size of training dataset

```{r}
plot_parameters(prior_posterior,carseat_train) 
plot_parameters(prior_posterior_small,carseat_train_small) 
plot_parameters(prior_posterior_tiny,carseat_train_tiny)

```


Posterior Prediction Check

```{r}
# Function that adds size of training dataset to pp_check
pp_check_info <- function(model) {
  pp_check(model) + ggtitle(str_c('n = ',as.character(nrow(model$data)))) +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5))
}

pp_check_info(stan_model)
pp_check_info(stan_model_small)
pp_check_info(stan_model_tiny)
```


Manually plot the outcome distribution to compare to the posterior check plot above 

```{r}
ggplot(aes(x=Sales),data=carseat_train) + geom_density() + theme_minimal()
```

Make predictions using the posterior distribution

```{r}
post_pred <- posterior_predict(stan_model,new_data = carseat_test,draws = 1000) %>%
  as_tibble()
```

```{r}
mcmc_areas(stan_model,pars=c('Advertising','Price')) + theme_bw()
mcmc_intervals(stan_model,pars=c('Advertising','Price')) + theme_bw()
posterior_vs_prior(stan_model)
```



Look at the posterior prediction distribution for a single observation

```{r}

row_num <- quo(`25`)
ggplot(aes(x=!!row_num),data=post_pred) + geom_density() + theme_minimal()

# Take a look at that same row number
print(carseat_test %>% select(Sales, Advertising, Price) %>% slice(as.numeric(as_label(row_num))))
```

