---
title: "Bayesian Modeling"
author: "Jesse Cambon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
    toc: true
---

```{r knit-settings, include=FALSE}
library(here)
source(here("rmd_config.R"))
```

References:
* http://appliedpredictivemodeling.com/data
* http://faculty.marshall.usc.edu/gareth-james/ISL/data.html

Todo:
* HDI
* Sigma Term
* References

## Setup

```{r,message=F,warning=F}
#library(AppliedPredictiveModeling) # datasets
library(ISLR) # datasets
library(skimr)
library(tidyverse)
library(wesanderson)
library(rstanarm)
library(bayestestR)
library(insight)
library(bayesplot)
library(broom)
library(rsample)
library(jcolors)
library(patchwork)
library(ggrepel)
library(knitr)

num_cores <-  parallel::detectCores()
options(mc.cores = num_cores)

set.seed(42) # for reproducibility
```

## Set input data and formula

Datasets and formulas:
* ISLR::Carseats : Sales ~ Advertising + Price
* ISLR::Credit : Limit ~ Income + Rating
* chickwts: weight ~ feed

```{r}
### Set input dataset here ################
split <- initial_split(chickwts, prop = 0.9)
############################################

### Set model equation here ##########################
model_formula = as.formula(weight ~ feed)
######################################################
```

```{r}
chickwts %>% group_by(feed) %>%
  summarize(n=n(),
            min=min(weight),
            median=median(weight),
            mean=mean(weight),
            max=max(weight)) %>%
            ungroup() %>%
  kable()
```


```{r}
ggplot(data=chickwts,aes(x=weight,fill=feed)) + 
  facet_wrap( ~ feed) +
  theme_minimal() +
  theme(legend.position='none') +
  geom_density(alpha=0.7)+
  scale_color_jcolors('default') +
  xlab('Weight') + ylab('')
```


C/V split
```{r}
train <- training(split) %>% as_tibble()
test  <- testing(split) %>% as_tibble()

train_small <- train %>% sample_n(30)
train_tiny <- train %>% sample_n(15)
```

Fit models

```{r}
lm_model <- glm(model_formula, data = train)
stan_model <- stan_glm(model_formula, data = train)
stan_model_small <- stan_glm(model_formula, data = train_small)
stan_model_tiny <- stan_glm(model_formula, data = train_tiny)
```

Extract posterior

http://mc-stan.org/rstanarm/reference/as.matrix.stanreg.html
```{r}
post1 <- as.data.frame(stan_model) %>% as_tibble()
```

Posterior Intervals

https://mc-stan.org/rstanarm/reference/posterior_interval.stanreg.html
https://mc-stan.org/rstanarm/articles/rstanarm.html
```{r}
rstanarm::posterior_interval(stan_model) %>% as.data.frame() %>%
  rownames_to_column('feed')
```

```{r}
tidy(lm_model,conf.int=T) %>% select(-std.error)
```




```{r}
# this could take a bit to run
post_descr <- describe_posterior(stan_model, test = c("p_direction","rope","bayesfactor"))

kable(post_descr)
```



Rope

```{r}
rope(stan_model)
rope(stan_model_small)
```



Markov Chain Diagnostics

```{r}
mcmc_trace(stan_model)
mcmc_trace(stan_model_small)
#mcmc_trace(stan_model_tiny)
```

Highest Density Intervals

```{r}
hdi(stan_model)
stan_model$coefficients
hdi(stan_model_small)
hdi(stan_model_tiny)
```

What percentage of each posterior distribution is greater than a certain cutoff value?

```{r}

cutoff_value <- -100 # define the cutoff

posterior <- get_parameters(stan_model,iterations=10000) %>% 
  pivot_longer(everything(),names_to='Parameter')

post_pct <- 
  posterior %>% filter(Parameter != '(Intercept)') %>%
  mutate(above=case_when(value > cutoff_value ~ 1, TRUE ~ 0)) %>%
  group_by(Parameter) %>%
  summarize(above_pct=mean(above)) %>%
  ungroup()

post_pct %>% kable()
```

### Posterior Distributions

```{r}
# Function that adds size of training dataset to mcmc_areas
mcmc_areas_info <- function(model,variables) {
  predictor_vars <- str_c('feed',unlist(stan_model$xlevels,use.names=F)[-1])
  
  mcmc_areas(model,pars=predictor_vars) + ggtitle(str_c('n = ',as.character(nrow(model$data)))) +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5))
}

mcmc_areas_info(stan_model,predictors)
mcmc_areas_info(stan_model_small,predictors)
mcmc_areas_info(stan_model_tiny,predictors) 
#mcmc_intervals(stan_model,pars=predictors) + theme_bw()
#posterior_vs_prior(stan_model)
```




```{r}
prior_summary(stan_model)
```

Draw from the prior and posterior distributions

```{r}
# Function for simulating prior and posterior distributions from stan model
sim_post_prior <- function(model) {
  # Simulate prior with bayestestR package
  prior <- simulate_prior(model) %>%
  pivot_longer(everything(),names_to='Parameter')

  # Simulate Posterior with insight package
  posterior <- get_parameters(model,iterations=10000) %>% 
  pivot_longer(everything(),names_to='Parameter')

  # Combine into one dataset
  combined <- prior %>% mutate(Distribution='Prior') %>% 
  bind_rows(posterior %>% mutate(Distribution='Posterior'))
  
  return(combined)
}

prior_posterior <- sim_post_prior(stan_model)
prior_posterior_small <- sim_post_prior(stan_model_small)
prior_posterior_tiny <- sim_post_prior(stan_model_tiny)

```

Plot our parameter prior and posterior distributions

```{r}

# Find the x,y coordinates for peak density in a sample
find_peak_density <- function(x_sample) {
  density_x <- density(x_sample)
  # Find coordinates for peak density
  x_max <- density_x$x[which.max(density_x$y)]
  y_max <- max(density_x$y)
  
  return(tibble(x=x_max,y=y_max))
}

# Function for plotting 
plot_parameters <- function(distribution_sample,train_data,plot_peaks=FALSE) {
    
  # data to plot - exclude intercept term
  plot_data <- distribution_sample %>% filter(!str_detect(Parameter,'Intercept'))
    
  # Points for labeling max density 
  # based loosely on: https://stackoverflow.com/questions/56520287/how-to-add-label-to-each-geom-density-line)
 density_coordinates <- plot_data %>% 
  group_by(Distribution,Parameter) %>%
  do(find_peak_density(.$value))
    
  base_plot <- ggplot(data=plot_data,
         aes(x=value,fill=Parameter)) +
    facet_wrap(~fct_rev(Distribution),scales='free') +
    theme_minimal() +
    scale_y_continuous(expand =c(0,0,0.15,0)) + # add spacing for labels
    geom_vline(xintercept=0,color='red',size=0.25,linetype='dashed') +
    theme(legend.position='top',
          legend.title=element_blank(),
          plot.title = element_text(hjust = 0.5)) +
    geom_density(alpha=0.4,size=0.05) + ggtitle(str_c('n = ',as.character(nrow(train_data)))) +
    xlab('') + ylab('') + scale_fill_jcolors('pal6') + 
    guides(color = guide_legend(reverse=T))
  
  if (plot_peaks == TRUE) {
    return(base_plot +
    geom_point(data=density_coordinates, aes(x=x, y=y),show.legend = F) +
    geom_text_repel(data=density_coordinates, aes(label=round(x,2),x=x, y=y),
                     force=1.5,size=4,show.legend = F))
  }
  else {
    return(base_plot)
  }
}
```

Compare parameter distributions by sample size of training dataset

```{r}
plot_parameters(prior_posterior,train) 
plot_parameters(prior_posterior_small,train_small) 
plot_parameters(prior_posterior_tiny,train_tiny)

```


## Predictive Posterior Distribution

```{r}
# Function that adds size of training dataset to pp_check
pp_check_info <- function(model) {
  pp_check(model) + ggtitle(str_c('n = ',as.character(nrow(model$data)))) +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5))
}

pp_check_info(stan_model)
pp_check_info(stan_model_small)
pp_check_info(stan_model_tiny)
```


Manually plot the outcome distribution to compare to the posterior check plot above 




```{r}
# Extract variables from formula
all_model_vars <- all.vars(model_formula)
outcome_var <- sym(all_model_vars[1])
predictors <- all_model_vars[-1]

ggplot(aes(x=!!outcome_var),data=train) + geom_density() + theme_minimal()
```

Make predictions using the posterior distribution

```{r}
post_pred <- posterior_predict(stan_model,new_data = test,draws = 1000) %>%
  as_tibble()
```


Look at the posterior prediction distribution for a single observation

```{r}
row_num <- quo(`2`)

true_value <- test %>% slice(as.numeric(as_label(row_num))) %>%
  pull(outcome_var)

ggplot(aes(x=!!row_num),data=post_pred) + geom_density() + theme_minimal() +
  geom_vline(xintercept=true_value,color='steelblue')

# Take a look at that same row number
print(test %>% slice(as.numeric(as_label(row_num))))
```

