---
title: "R Quickstart"
author: "Jesse Cambon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
    toc: true
---

```{r knit-settings, include=FALSE}
library(here)
source(here("rmd_config.R"))
```

Goal: Simple, minimal code for getting started with R.

## Todo 
* Clearly label before and after datasets to illustrate data transformations
* basic minimal ggplots (histogram, point, bar, line)

## Setup

```{r setup, message=FALSE, results=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)

# Set default ggplot theme
theme_set(theme_bw()+
  theme(legend.position = "top",
            plot.subtitle= element_text(face="bold",hjust=0.5),
            plot.title = element_text(lineheight=1, face="bold",hjust = 0.5)))
```

## Data Manipulation

### Stacking Data

Initial 'mpg' Dataset:
```{r,echo=F} 
kable(head(mpg,3))
```

```{r}
mpg1 <- mpg %>% slice(1:2) %>% 
  select(manufacturer,model,hwy,cty) %>%
  mutate(dataset=1)

mpg2 <- mpg %>% slice(44:45) %>%
  select(manufacturer,model,hwy,cty) %>%
  mutate(dataset=2)

mpg_stack_vert <- mpg1 %>% 
  bind_rows(mpg2)

mpg3 <- mpg %>% slice(1:2,5:6) %>%
  select(displ,year)

mpg_stack_horz <- mpg_stack_vert %>%
  bind_cols(mpg3)

car_type <- mpg %>% select(manufacturer,model,class) %>%
  distinct() # distinct rows only

joined <- mpg_stack_horz %>%
  left_join(car_type,by=c('manufacturer','model')) %>% 
  select(-dataset,everything())
```

### Long to Wide
Initial Data:
```{r,echo=F}
kable(head(us_rent_income,4))
```

* pivot_wider
  * names_from: column containing values that we will use for our new column names
  
```{r}
col_ratio <- us_rent_income %>%
  select(-GEOID,-moe) %>%
  pivot_wider(names_from = variable, values_from = estimate) %>% 
  drop_na() %>%   # drop missing values
  mutate(income_rent_ratio = income / (12*rent))
```

Income and Rent are now in separate columns:

```{r,echo=F}
kable(head(col_ratio,4))
```

### Wide to Long 
Initial Data:
```{r,echo=F}
kable(head(world_bank_pop,3))
```

* pivot_longer
  * cols (1st arg): what columns do we want to pivot? (ie. subtract ones we don't want to)
  * names_to : the name of new column holding the column names as values
  * values_to : name of new column containing values
* seq(start, stop, increment)  ->  generates sequence

```{r} 
wb_pop <- world_bank_pop %>%
  pivot_longer(c(-country,-indicator), names_to = "year", values_to = "value") %>%
  mutate(year=as.numeric(year)) %>% # convert to numeric
  filter(year %in% seq(2000,2016,2))
```

After:
```{r,echo=F} 
kable(head(wb_pop,3))
```


### Counting
```{r}
count_cyl <- mpg %>%
  count(cyl)
```

```{r,echo=F}
kable(count_cyl)
```

### Calculate Summary Stats
```{r}
mpg_stats <- mpg %>% select(class,hwy) %>%
  mutate(class_c=case_when(class %in% c("2seater","subcompact") ~ "subcompact",
                               TRUE ~ class)) %>%
  group_by(class_c) %>%
  summarize(count=n(),
            max_hwy=max(hwy),
            min_hwy=min(hwy),
            median_hwy=median(hwy),
            mean_hwy=mean(hwy)) %>%
  ungroup() %>%
  arrange(desc(count)) # sort dataset
```

Note that '2seater' is reclassified as 'subcompact'

```{r,echo=F}
kable(mpg_stats)
```

## Visualizations

### Bar Chart

* use fill argument in ggplot() to set bar color based on a variable
* reorder() orders the bars
```{r}
# A simple bar chart - average heights of the species
# the reorder command orders our bars in order of descending height
ggplot(data=mpg_stats,
    aes(x = reorder(class_c,-mean_hwy), y=mean_hwy)) +
geom_bar(stat='identity',position='dodge',color='black') +
scale_y_continuous(expand = c(0,0,0.08,0)) +    # plot margins
geom_text(aes(label=round(mean_hwy)), vjust=-0.5) +  # labelling
theme(legend.position="none", # no legend (in case we want to use fill)
      panel.grid = element_blank()) + # turn off grid
labs(title='') +
xlab('') +
ylab('')
```


```{r histogram, eval=FALSE}
# Histogram with autobinning based on gender
ggplot(mpg,aes(hwy)) +
geom_histogram(aes(fill=cyl)) +
scale_y_continuous(expand = c(0,0,0.08,0)) + 
xlab('') + ylab('')
```

## Line

```{r line}
ggplot(wb_pop %>% filter(country %in% c("USA","CAN","MEX") & indicator == "SP.POP.GROW"),
          aes(x=year,y=value,color = country)) +
  theme_classic() +
geom_line() + geom_point() + # lines and points
scale_x_continuous(expand=c(0.02,0,0.02,0)) +
#scale_y_continuous(labels=scales::comma) + 
labs(title='',
     caption='') +
theme(legend.title = element_blank(),
      panel.grid.minor.x = element_blank(),
      legend.text=element_text(size=10),
      legend.position='right') +
xlab('Year') +
ylab('Population Growth (millions)') +
# make legend lines bigger
guides(colour = guide_legend(override.aes = list(size=2.5))) 

```


## Lollipop

```{r lollipop}
  ggplot(data=col_ratio %>% arrange(desc(rent)) %>% head(8), aes(x=NAME, y=rent) ) +
    geom_segment( aes(x=reorder(NAME,rent) ,xend=NAME, y=0, yend=rent), color="grey") +
    geom_point(size=3, color="navy") +
   theme_minimal() +
  theme(    plot.subtitle= element_text(face="bold",hjust=0.5),
            plot.title = element_text(lineheight=1, face="bold",hjust = 0.5),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.x = element_blank(),
      legend.position="none"
    ) +
  coord_flip() +
    # expand gets rid of space between labels stem of lollipops
    scale_y_continuous(expand = c(0, 0,0,70)) + 
    labs(title='States With Highest Rent',
        caption='') +
    xlab("") +
    ylab('')
```


