---
title: "Survival Models"
author: "Jesse Cambon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
    toc: true
---

Survival Analysis 

* Kaplan-Meier Plots
* Log-rank test
* Cox Proportional Hazard Model
* Parametric survival models
* Bayesian Approaches

```{r knit-settings, include=FALSE}
library(here)
source(here("rmd_config.R"))
```

Reference:

https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html


```{r setup, message = F, warning = F}
library(survival)
library(survminer)
library(tidyverse)
library(broom)
```


# Kaplan-Meier 

status: censoring status 1=censored, 2=dead. Can also use TRUE/FALSE see documentation for
event in ?Surv

```{r}
lung %>% count(status)
```


```{r}
ggsurvplot(
    fit = survfit(Surv(time, status) ~ sex, data = lung), 
    xlab = "Days", 
    ylab = "Survival Probability")$plot
```

# Log-Rank Test

Test if there was a statistically significant difference in survival time between the groups

```{r}
survdiff(Surv(time, status) ~ sex, 
         data = lung, 
         rho = 0 # log-rank, see ?survdiff
         )
```

# Cox Proportional Hazard Model

* Multivariate "semi-parametric" regression approach
* Assumes hazard can change over time, but is proportional between groups at all points in time (ie. hazard ratio is constant over time).


```{r}
cox_fit <- coxph(Surv(time, status) ~ sex + age + ph.ecog,
        data = lung)

# Exponentiate coefficients to get hazard ratios
cox_hr <- tidy(cox_fit, exponentiate = TRUE, conf.int = TRUE)
```

Survival curve

```{r}
ggsurvplot(survfit(cox_fit), data = lung, risk.table = TRUE)
```

Plot Hazard Ratios

```{r}
ggplot(data=cox_hr,
          aes(x = term, y = estimate)) +
geom_point() +
scale_y_continuous() +
geom_hline(yintercept=0,color='grey') +
coord_flip() +
  theme_bw() +
  theme(plot.title = element_text(lineheight = 1, face="bold",hjust = 0.5)) +
geom_pointrange(mapping = aes(ymin = conf.low, ymax = conf.high)) + 
xlab('Term') + ylab('HR') + geom_hline(yintercept = 1, color = "grey")
```

### Predictions 

```{r}
sample_obs <- lung %>% 
  sample_n(2, seed = 104) %>%
  mutate(id = 1:n()) %>%
  select(id, status, everything())

cox_pred <- predict(cox_fit, newdata = sample_obs, type = 'expected')

```



### Validation

Reference: 
http://www.sthda.com/english/wiki/cox-model-assumptions 


```{r}
concordance(cox_fit)
```


Look at residuals

```{r}
ggcoxdiagnostics(cox_fit, type = "deviance", ox.scale = 'observation.id')
ggcoxdiagnostics(cox_fit, type = "deviance", ox.scale = 'linear.predictions')
```

```{r}
ggcoxdiagnostics(cox_fit, type = 'dfbeta')
```



Test proportional hazards assumption

```{r, fig.height = 7, fig.wdith = 5}
zph_fit <- cox.zph(cox_fit)
ggcoxzph(zph_fit)
```

## Parametric Survival Model

Accelerated Failure Time models, an alternative to cox regression

```{r}
aft_fit <- survreg(Surv(time, status) ~ sex + age + ph.ecog,
        dist = 'weibull',
        data = lung)

# Exponentiate coefficients to get hazard ratios
aft_hr <- tidy(aft_fit, exponentiate = TRUE, conf.int = TRUE)

aft_hr
```





## Bayesian Survival Models

```{r}
library(brms)
library(bayesplot)
options(mc.cores = parallel::detectCores())
```

- http://paul-buerkner.github.io/brms/reference/kidney.html
- https://mc-stan.org/rstanarm/reference/adapt_delta.html 

```{r, message = FALSE,  warning = F, error = F, results = 'hide'}
# fit weibull model
fit2 <- brm(time | cens(censored) ~ sex + disease + (1 | patient), 
            data = kidney, family = weibull(), cores = 4, 
            control = list(adapt_delta = 0.97))
```


```{r}
summary(fit2)

prior_summary(fit2)
```



```{r}
pp_check(fit2)

pp_check(fit2, type = 'intervals')
```
```{r, fig.height =  8, fig.width = 4}
mcmc_areas(fit2,  regex_pars = c('b_*', 'r_*'))
```

https://mc-stan.org/bayesplot/reference/PPC-censoring.html

```{r}
yrep <- posterior_predict(fit2)
```


```{r}
ppc_km_overlay(kidney$time, yrep, status_y = kidney$censored)
```

